---
title: "Baby Names Analysis"
author: "Milan Bertolutti"
date: 2026-09-10
---

Data: [babynames](https://cran.r-project.org/package=babynames), from [US Social Security Administration](https://www.ssa.gov/oact/babynames/limits.html) records. Licence: CC0.

```{r}
#| message: false
#| warning: false
#setup-chunk
library(babynames)
library(dplyr)
library(ggplot2)
```

We will start our analysis by looking at the data set.

```{r}
babynames
```

Let's start by viewing some of the most popular names scaled with the number of babies born each year (prop).

```{r}
baby_names_sorted <- babynames |> arrange(desc(prop))

head(baby_names_sorted, 10)
```

Let's now check out how popular my name has been over the years, scaled with birth rate of course:

```{r}
my_name_sorted <- babynames |> filter(tolower(name) == "milan") |> arrange(desc(prop))

my_name_sorted
```

It looks like my name is getting more popular, and used to be more popular than it was at the end of the 20th century! Talk about being a trend setter in this new age.

I have always wondered if my name has been more popular as a female or a male name. Let's get to the bottom of this. 

We will start by visualizing the data.

```{r}
#| label: fig-milan-by-sex
#| fig-cap: "Babies named Milan per million, by sex."
#plot code generated by Claude
library(ggplot2)

# add every year for each sex so years with no data show as gaps instead of a joined line
all_years <- expand.grid(year = 1880:2017, sex = c("F", "M"), stringsAsFactors = FALSE)

milan_by_year <- all_years |>
  left_join(my_name_sorted, by = c("year", "sex")) |>
  mutate(per_million = prop * 1e6)

ggplot(milan_by_year, aes(x = year, y = per_million, colour = sex)) +
  geom_line(linewidth = 0.8, na.rm = TRUE) +
  scale_colour_manual(values = c(F = "#eb6834", M = "#2a78d6"),
                      labels = c(F = "Female", M = "Male")) +
  labs(title = "Babies named Milan per million, by sex",
       x = "Year", y = "Milans per million babies of that sex", colour = NULL) +
  theme_minimal() +
  theme(legend.position = "top")

```

By looking at the visualization, we can determine that over the last century Milan has been more male dominant. Although, looking at the tail of the graph there was a decade where it was significantly more popular for females. 

This code below shows the total number of recorded male and female Milans. Male Milans are a little more dominant when compared to the female birth rate.

```{r}
my_name_by_sex <- my_name_sorted |> group_by(sex) |> summarise(total_per_sex = sum(n), average_prop = mean(prop))

my_name_by_sex
```

Since I was born in the early 2000s, let's zoom the graph in and get more insight on my name over the 21st century. 

```{r}
#| label: fig-milan-recent
#| fig-cap: "Babies named Milan per million, by sex, 1997 to 2017."
# add every year for each sex so years with no data show as gaps instead of a joined line
all_years <- expand.grid(year = 1997:2017, sex = c("F", "M"), stringsAsFactors = FALSE)

milan_by_year <- all_years |>
  left_join(my_name_sorted, by = c("year", "sex")) |>
  mutate(per_million = prop * 1e6)

ggplot(milan_by_year, aes(x = year, y = per_million, colour = sex)) +
  geom_line(linewidth = 0.8, na.rm = TRUE) +
  scale_colour_manual(values = c(F = "#eb6834", M = "#2a78d6"),
                      labels = c(F = "Female", M = "Male")) +
  labs(title = "Babies named Milan per million, by sex over the last 20 years recorded",
       x = "Year", y = "Milans per million babies of that sex", colour = NULL) +
  theme_minimal() +
  theme(legend.position = "top")
```

So, looking at it now, we can see that with people around my age or younger, there has been a higher volume of female Milans.