---
title: "Baby Names Analysis"
author: "Milan Bertolutti"
date: 2026-09-10
---

Data: [babynames](https://cran.r-project.org/package=babynames), from [US Social Security Administration](https://www.ssa.gov/oact/babynames/limits.html) records. Licence: CC0.

```{r}
#setup-chunk
library(babynames)
library(dplyr)
```

We will start are analysis by looking at the data set.

```{r}
babynames
```

Lets start by viewing some of the most popular names scaled with the amount of baby born each year (prop)

```{r}
baby_names_sorted <- babynames |> arrange(desc(prop))

head(baby_names_sorted, 10)
```

Lets now check out how popular my name has been over the years, scaled with birth rate of course:

```{r}
my_name_sorted <- babynames |> filter(tolower(name) == "milan") |> arrange(desc(prop))

my_name_sorted
```

It looks like my name is getting more popular, and used to be more popular then it was at the end of the 20th century! Talk about being a trend setter in this new age.

I have always wondered if my name has been more popular as a female or a male name. Lets get to the bottom of this. 

```{r}
my_name_by_sex <- my_name_sorted |> group_by(sex) |> summarise(total_per_sex = sum(n))

my_name_by_sex
```

